\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-12 smurf_paper.tex
add ref for LABOCA map-maker, more acknowledgements

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar \rms\ may be calculated for all of the data points from
a given bolometer to characterise its noise distribution (since the
noise at any instant in time is uncorrelated with others, and with the
data from other bolometers), greatly simplifying the measurement of
noise properties and the inversion step that is so complicated in the
maximum-likelihood methods. In principle, it should also be possible
to recover a wider dynamic range in angular scales than possible for
the simplest direct methods, with an increase in calculation time that
is merely linear in the number of iterations needed, while using a
comparable amount of memory. The iterative method of
\citet{wright1996} specific to chopped (CMB) data was applied to SCUBA
observations by \citet{johnstone2000}, although in this case the
chopping removes most of the low-frequency noise, and greatly
simplifies the problem (at the expensive of large-scale structure in
the data). More similar to the algorithm described here was the
iterative pipeline recipe for fitting and removing baseline drifts in
SCUBA scan-map data as the astronomical image estimate improved
\citep{1999ASPC..172..171J}. The two closest modern relatives of SMURF
are the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep[][which uses iterative PCA
cleaning to avoid the suppression of large-scale
structure]{aguirre2011}.

@@

Finally, we experimented with an alternative to \model{COM} for
removing correlated atmospheric noise. Rather than subtracting the
average signal at each time slice, \model{PLN} fits a plane to the
observed signal at each instant. We found that there was no obvious
improvement (either in terms of reducing the $1/f$ knee or reducing
the noise in final maps). This result is basically consistent with
those from earlier instruments, and suggests that the angular scale of
the emitting regions in the atmosphere are un-resolved at the \scuba\
focal plane. \citet{sayers2010}, in particular, examined several
different ways of modelling and removing the correlated atmospheric
noise from Bolocam data at 143 and 268\,GHz (also atop Mauna Kea),
finding only marginal improvements by fitting a plane, or even using
higher-order polynomials (see their figure~10, as well as the
discussion and references to earlier work in their section~4.3). In
\citet{aguirre2011} iterative PCA cleaning was used to remove the
atmosphere since the simpler methods mentioned left substantial
residuals in their data (similar to the \scuba\ data described here,
although PCA cleaning is prohibitively slow in this case).

@@

This paper has described the Submillimetre User Reduction Facility
(SMURF), which was designed to produce maps from the rapidly sampled
$\sim10^4$ bolometers of the \scuba\ instrument. While similar to
other algorithms in the literature that iteratively model and remove
correlated noise components from the data, successively improving the
map \citep[e.g.,][]{kovacs2008,aguirre2011}, we have shown that a
relatively simpler approach provides good results for \scuba\ data,
with reasonable computational requirements. This conclusion is
encouraging for the development of future large-format bolometer
arrays, for which our specific approach should be useful when even
larger volumes of data are involved.

@@

The main problem with the iterative solution is divergence on large
angular scales, due to the degeneracy between the map, and the
low-frequency signal components that are removed. A simple strategy of
constraining empty regions of the map to zero (using either a
user-supplied mask for known sources, or an iterative determination of
signal above some \snr\ threshold) provides good constraints for
compact objects, and bright/extended structures, allowing us to
effectively recover angular scales up to the order of the array
footprint (approximately 5\,arcmin).

@@

The James Clerk Maxwell Telescope is operated by the Joint Astronomy
Centre on behalf of the Science and Technology Facilities Council of
the United Kingdom, the Netherlands Organisation for Scientific
Research, and the National Research Council of Canada. Additional
funds for the contruction of \scuba\ were provided by the Canada
Foundation for Innovation. EC thanks CANARIE/CANFAR for additional
funding. This research was supported in part by the Natural Sciences
and Engineering Research Council of Canada. We thank Mark Halpern,
Matthew Hasselfield, and Gaelen Marsden for many useful discussions.
The authors thank the members of the SCUBA-2 commissioning team for
testing the map-maker and reporting anomalies; in particular Antonio
Chrysostomou, Jessica Dempsey, and Wayne Holland. We acknowledge the
contributions of Dennis Kelly, Alex van Engelen and Jennifer Balfour
for early investigations related to SMURF; and Mark Thompson, Craig
Walther and S\'{e}verin Gaudet for being on the Critical Design Review
panel. Finally, we thank Per Friberg and Gary Davis for their helpful
comments on the manuscript.

>>>>

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar \rms\ may be calculated for all of the data points from
a given bolometer to characterise its noise distribution (since the
noise at any instant in time is uncorrelated with others, and with the
data from other bolometers), greatly simplifying the measurement of
noise properties and the inversion step that is so complicated in the
maximum-likelihood methods. In principle, it should also be possible
to recover a wider dynamic range in angular scales than possible for
the simplest direct methods, with an increase in calculation time that
is merely linear in the number of iterations needed, while using a
comparable amount of memory. The iterative method of
\citet{wright1996} specific to chopped (CMB) data was applied to SCUBA
observations by \citet{johnstone2000}, although in this case the
chopping removes most of the low-frequency noise, and greatly
simplifies the problem (at the expensive of large-scale structure in
the data). More similar to the algorithm described here was the
iterative pipeline recipe for fitting and removing baseline drifts in
SCUBA scan-map data as the astronomical image estimate improved
\citep{1999ASPC..172..171J}. The closest modern relatives of SMURF are
the Comprehensive Reduction Utility \citep[CRUSH,][]{kovacs2008} for
the Submillimeter High-Angular Resolution Camera 2 (SHARC-2), the
Bolometer array data Analysis software (BoA) for the Large APEX
BOlometer CAmera \citep[LABOCA,][]{siringo2009}, and the pipeline
developed for the Bolocam Galactic Plane Survey \citep[][which uses
iterative PCA cleaning to avoid the suppression of large-scale
structure]{aguirre2011}.

@@

Finally, we experimented with an alternative to \model{COM} for
removing correlated atmospheric noise. Rather than subtracting the
average signal at each time slice, \model{PLN} fits a plane to the
observed signal at each instant. We found that there was no obvious
improvement (either in terms of reducing the $1/f$ knee or reducing
the noise in final maps). This result is basically consistent with
those from earlier instruments, and suggests that the angular scale of
the emitting regions in the atmosphere are un-resolved at the \scuba\
focal plane. \citet{sayers2010}, in particular, examined several
different ways of modelling and removing the correlated atmospheric
noise from Bolocam data at 143 and 268\,GHz (also atop Mauna Kea),
finding only marginal improvements by fitting a plane, or even using
higher-order polynomials (see their figure~10, as well as the
discussion and references to earlier work in their section~4.3). In
\citet{aguirre2011} iterative PCA cleaning was used to remove the
atmosphere since the simpler methods mentioned left substantial
residuals in their data (similar to the \scuba\ data described here,
although PCA cleaning is prohibitively slow in our case).

@@

This paper has described the Submillimetre User Reduction Facility
(SMURF), which was designed to produce maps from the rapidly sampled
$\sim10^4$ bolometers of the \scuba\ instrument. While similar to
other algorithms in the literature that iteratively model and remove
correlated noise components from the data, successively improving the
map \citep[e.g.,][]{kovacs2008,siringo2009,aguirre2011}, we have shown
that a relatively simpler approach provides good results for \scuba\
data, with reasonable computational requirements. This conclusion is
encouraging for the development of future large-format bolometer
arrays, for which our specific approach should be useful when even
larger volumes of data are involved.

@@

The main problem with the iterative solution is divergence on large
angular scales, due to the degeneracy between the map, and the
low-frequency signal components that are removed. A simple strategy of
constraining empty regions of the map to zero (using either a
user-supplied mask for known sources, or an iterative determination of
signal above some \snr\ threshold) provides good constraints for both
compact objects, and bright/extended structures, allowing us to
effectively recover angular scales up to the order of the array
footprint (approximately 5\,arcmin).

@@

The James Clerk Maxwell Telescope is operated by the Joint Astronomy
Centre on behalf of the Science and Technology Facilities Council of
the United Kingdom, the Netherlands Organisation for Scientific
Research, and the National Research Council of Canada. Additional
funds for the contruction of \scuba\ were provided by the Canada
Foundation for Innovation.  This research was supported in part by the
Natural Sciences and Engineering Research Council of Canada.  EC
thanks CANARIE/CANFAR for additional funding.  The authors thank the
members of the SCUBA-2 commissioning team for testing the map-maker
and reporting anomalies; in particular Antonio Chrysostomou, Jessica
Dempsey, and Wayne Holland.  We also Mark Halpern, Matthew
Hasselfield, and Gaelen Marsden for many useful discussions, and
observers who provided useful feedback; especially David Nutter and
Todd McKenzie.  We acknowledge the contributions of Dennis Kelly, Alex
van Engelen and Jennifer Balfour for early investigations related to
SMURF; and Mark Thompson, Craig Walther and S\'{e}verin Gaudet for
being on the Critical Design Review panel. Finally, we thank Per
Friberg and Gary Davis for their helpful comments on the manuscript.
